EuPPollNet: A European database of plant-pollinator networks

### 2024-04-10 Manuscript/2_Figure_2.Rmd
results, sampling coverage

```{r echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE, out.width="95%", fig.cap= "\\textbf{Figure 2}.", fig.width=10, fig.height=6}

#Load libraries
library(dplyr)
library(ggplot2)
library(scales)
library(patchwork)

@@

#Plot
o1 = ggplot(poll_spread, 
aes(reorder(Pollinator_accepted_name,-Percent_total), Percent_total)) +
geom_bar(stat = "identity", 
           color = "#D55E00",
           lwd = 0.03, fill = "#D55E00") +
ylab("Shared pollinators (%)") +
theme_minimal()+
theme(axis.text.x = element_blank(),
  axis.ticks.x=element_blank(),
  axis.ticks.y = element_line(size = 0.2),
  panel.grid.major.x = element_blank(),
  panel.grid.minor.y = element_blank(),
  panel.grid.major.y = element_line(size=0.2, linetype = 2)) +
xlab("Pollinator species") +
ylab("Occurrence across sampling sites") + 
geom_vline(xintercept = quantile_50, linetype="dashed", 
color = "black", size=0.35) + 
annotate(geom="text", x=1450, y=0.60, label="50th percentile",color="black", size = 3) +
annotate(geom="text", x=140, y=0.67, label=most_common_poll[1],color="black",  fontface = 'italic', size = 3)+
annotate(geom="text", x=200, y=-0.05, label="Common spp."
      ,color="black",  size = 3) +
annotate(geom="text", x=1800, y=-0.05, label="\'Rare\' spp."
      ,color="black",  size = 3)  +
scale_y_continuous(breaks = seq(0, 0.7, by= 0.25), expand = c(0, 0)) + 
coord_cartesian(xlim=c(0, nlevels(factor(poll_spread$Pollinator_accepted_name))),
                ylim=c(0,0.7), clip="off") +
ggtitle("(d)")

@@

o2 = ggplot(plant_spread, 
aes(reorder(Plant_accepted_name,-Percent_total), Percent_total, fill = Plant_accepted_name)) +
geom_bar(stat = "identity", 
           color = "#009E73",
           lwd = 0.03, fill = "#009E73") +
theme_minimal()+
theme(axis.text.x = element_blank(),
         axis.ticks.x=element_blank(),
         axis.ticks.y = element_line(size = 0.2),
         panel.grid.major.x = element_blank(),
         panel.grid.minor.y = element_blank(),
         panel.grid.major.y = element_line(size=0.2, linetype =  2)) +
xlab("Plant species") +
ylab("Occurrence across sampling sites") + 
geom_vline(xintercept = quantile_50, linetype="dashed", 
          color = "black", size=0.35) + 
annotate(geom="text", x=950, y=0.60, label="50th percentile",color="black", size = 3) + 
annotate(geom="text", x=100, y=0.4, label=most_common_plant[1]
      ,color="black",  fontface = 'italic', size = 3)+
annotate(geom="text", x=150, y=-0.05, label="Common spp."
      ,color="black",  size = 3) +
annotate(geom="text", x=1200, y=-0.05, label="\'Rare\' spp."
      ,color="black",  size = 3) + 
scale_y_continuous(breaks = seq(0, 0.7, by= 0.25), expand = c(0, 0), labels = label_number(accuracy = 0.01)) +
coord_cartesian(xlim=c(0, nlevels(factor(plant_spread$Plant_accepted_name))),
                ylim=c(0, 0.7), clip="off") +
ggtitle("(e)")

plant_poll = readRDS("../Data/Working_files/collectors_curves_plant_ploll.rds")
rarefaction_curve1 = readRDS("../Data/Working_files/rarefaction_curve_plant_poll.rds")

#Plot curves
pp1 = plant_poll %>% 
ggplot(aes(x = Plant_sampled, y= Unique_spp_cumulative, group = iteration)) +
geom_line(color = "gray") +
geom_smooth(data = rarefaction_curve1, 
  aes(x = Plant_sampled, y = mean_curve), inherit.aes = FALSE,method = "loess", color = "black", size = 0.65, linetype = "solid", se = FALSE, span = 0.5) +  
geom_point(data = rarefaction_curve1, x = max(rarefaction_curve1$Plant_sampled), y = max(rarefaction_curve1$mean_curve), inherit.aes = FALSE, size=2) +
ylab("Pollinator species") +
xlab("Plant species") +
theme_bw() +
coord_cartesian(expand = FALSE, clip = "off") +
ggtitle("(f)") +
scale_x_continuous(breaks = seq(0, 1400, by = 400), limits=c(0,1400)) +
ylim(0,2500)

#Plot all
a = p1 + p2 + p3 + plot_layout(guides = "collect") & theme(legend.title=element_blank(),
      axis.title = element_text(size=12, face = "bold"),
      axis.text = element_text(size=10),
      plot.title = element_text(size=18))
b = o1 + o2 + pp1 & 
  theme(axis.title = element_text(size=12, face = "bold"),
        axis.text = element_text(size=10),
      plot.title = element_text(size=18))

a/b

>>>>

```{r Load libraries, echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE}
library(dplyr) #For handling fata 
library(sf) #For handling coordinates and plotting
library(scales) #For plotting (decimals on axes)
library(lubridate) #To operate with dates
library(tidyr) #Reshape data (wide format)
library(stringr) #Fix/modify strings
library(forcats)
library(readr)
#Load libraries for plotting
library(giscoR) #Maps
library(ggstar) #cool shapes as points
library(ggplot2) #general plotting
library(ggtree) #phylo trees
library(ggpattern) #cool shapes in barplots
library(patchwork) #plot organisation
library(cowplot) #plot organisation
library(viridis) #colours
library(ggnewscale)
library(ggtreeExtra)
library(kableExtra)

```{r echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE, out.width="95%", fig.cap= "\\textbf{Figure 2}. Graphs (a-b-c) indicate the accumulation curves for pollinators, plants, and their interactions across sampling sites. Grey solid lines represent 100 randomized accumulation curves, the black solid lines represents the interpolated curve (i.e., the mean across curves), and the red dashed lines illustrates the extrapolated curve for approximately 3000 sampling sites. The solid black points indicate the number of species and interactions contained in the database. Graph (d) shows the accumulation curve of pollinator species across increasing number of plant species. This last graph uses the same color and shape structure as the ones in the top panel. Graphs (e-f) indicate the percentage of occurrence (i.e., incidence) of plant and pollinator species across sampling sites. Species on the left (i.e., common) are found in many sampling sites, while species on the right (i.e., rare) are found in few or only a single sampling site.", fig.width=10, fig.height=6, out.width="115%"}

@@

#Plot
o1 = ggplot(poll_spread, 
aes(reorder(Pollinator_accepted_name,-Percent_total), Percent_total)) +
geom_bar(stat = "identity", 
           color = "#D55E00",
           lwd = 0.03, fill = "#D55E00") +
ylab("Shared pollinators (%)") +
theme_minimal()+
theme(axis.text.x = element_blank(),
  axis.ticks.x=element_blank(),
  axis.ticks.y = element_line(size = 0.2),
  panel.grid.major.x = element_blank(),
  panel.grid.minor.y = element_blank(),
  panel.grid.major.y = element_line(size=0.2, linetype = 2)) +
xlab("Pollinator species") +
ylab("Occurrence across sampling sites") + 
geom_vline(xintercept = quantile_50, linetype="dashed", 
color = "black", size=0.35) + 
annotate(geom="text", x=1450, y=0.60, label="50th percentile",color="black", size = 3) +
annotate(geom="text", x=140, y=0.67, label=most_common_poll[1],color="black",  fontface = 'italic', size = 3)+
annotate(geom="text", x=200, y=-0.05, label="Common spp."
      ,color="black",  size = 3) +
annotate(geom="text", x=1800, y=-0.05, label="\'Rare\' spp."
      ,color="black",  size = 3)  +
scale_y_continuous(breaks = seq(0, 0.7, by= 0.25), expand = c(0, 0)) + 
coord_cartesian(xlim=c(0, nlevels(factor(poll_spread$Pollinator_accepted_name))),
                ylim=c(0,0.7), clip="off") +
ggtitle("(e)")

@@

o2 = ggplot(plant_spread, 
aes(reorder(Plant_accepted_name,-Percent_total), Percent_total, fill = Plant_accepted_name)) +
geom_bar(stat = "identity", 
           color = "#009E73",
           lwd = 0.03, fill = "#009E73") +
theme_minimal()+
theme(axis.text.x = element_blank(),
         axis.ticks.x=element_blank(),
         axis.ticks.y = element_line(size = 0.2),
         panel.grid.major.x = element_blank(),
         panel.grid.minor.y = element_blank(),
         panel.grid.major.y = element_line(size=0.2, linetype =  2)) +
xlab("Plant species") +
ylab("Occurrence across sampling sites") + 
geom_vline(xintercept = quantile_50, linetype="dashed", 
          color = "black", size=0.35) + 
annotate(geom="text", x=950, y=0.60, label="50th percentile",color="black", size = 3) + 
annotate(geom="text", x=100, y=0.4, label=most_common_plant[1]
      ,color="black",  fontface = 'italic', size = 3)+
annotate(geom="text", x=150, y=-0.05, label="Common spp."
      ,color="black",  size = 3) +
annotate(geom="text", x=1200, y=-0.05, label="\'Rare\' spp."
      ,color="black",  size = 3) + 
scale_y_continuous(breaks = seq(0, 0.7, by= 0.25), expand = c(0, 0), labels = label_number(accuracy = 0.01)) +
coord_cartesian(xlim=c(0, nlevels(factor(plant_spread$Plant_accepted_name))),
                ylim=c(0, 0.7), clip="off") +
ggtitle("(f)")

#Pollinator by plant species
plant_poll_curves = readRDS("../Data/Working_files/collectors_curves_plant_ploll.rds")
plant_poll_output = readRDS("../Data/Working_files/plant_poll_spp_sampling_coverage.rds")

#Create dataset
d_pp = plant_poll_output$iNextEst$Plant_id
#Check cols
d_pp$col = d_pp$method

#Now create tibble with same structure
a_poll = tibble(t = plant_poll_curves$Plant_sampled, 
           method = as.character(plant_poll_curves$iteration), 
           qD = plant_poll_curves$Unique_spp_cumulative,
           col = "collectors")

#Bind datasets 
d_pp1 = bind_rows(d_pp, a_poll)
d_pp1_point = d_pp1 %>%  filter(method == "observed")
d_pp1_lines = d_pp1 %>%  filter(!method == "observed")

#Plot
#Order levels
d_pp1_lines$col = factor(d_pp1_lines$col, levels = c("collectors", "interpolated", "extrapolated"))
#Plot
pp1 = ggplot(d_pp1_lines, aes(x = t, y = qD, group = method, color = col)) +
geom_line(aes(linetype = col)) +
scale_colour_manual(values = c("gray", "black", "red"), labels = c("Randomizations", "Interpolated", "Extrapolated")) +
scale_linetype_manual(values = c("solid", "solid", "dashed"),  labels = c("Randomizations", "Interpolated", "Extrapolated")) + 
geom_ribbon(aes(ymin = qD.LCL, 
    ymax=qD.UCL, fill = col),alpha=0.2, colour = NA, show.legend = FALSE)+
scale_fill_manual(values = c("gray", "black", "red")) +
geom_point(data = d_pp1_point, col = "black") + 
ylab("Pollinator species") +
xlab("Plant species") +
theme_bw() +
coord_cartesian(expand = FALSE, clip = "off") +
ggtitle("(d)") +
scale_x_continuous(breaks = seq(0, 2800, by = 1000), limits=c(0,2800)) +
scale_y_continuous(breaks = seq(0, 2600, by = 1000), limits=c(0,2600)) +
theme(legend.position = "none")

#Plot all
p1 + p2 + p3 + pp1 + o1 + o2 + plot_layout(guides = "collect", nrow = 2, ncol=3) & theme(legend.title=element_blank(),
      axis.title = element_text(size=12, face = "bold"),
      axis.text = element_text(size=10),
      plot.title = element_text(size=18))

#a/b

### 2024-04-10 Manuscript/Ms.qmd
results, sampling coverage

**Motivation:** Pollinators play a crucial role in maintaining Earth's terrestrial biodiversity and human food production by mediating sexual reproduction for most flowering plants. However, their diversity and role as pollinators are increasingly compromised by rapid human-induced environmental changes. One of the major challenges for pollinator conservation, is the lack of robust generalisable data across space and time to comprehend the conservation status and population trends among different pollinator species. Here, we present the EuPPollNet database, a fully reproducible European-level database containing harmonized taxonomic data on plant-pollinator interactions referenced in both space and time. This database offers an open workflow that allows researchers to track decisions and edit them according to their preferences, while also providing other ecological variables of interest. We hope this database can help researchers to: 1) identify taxonomic, ecological and geographical gaps of knowledge on plant-pollinator interactions; and 2) explore the impacts of global change to guide future conservation planning for both plant and pollinator species.

@@

**KEYWORDS**\
\begingroup
\small
plant-pollinator interactions, flowering plants, Angiosperms, pollinators, Hymenoptera, Diptera, Lepidoptera, Coleoptera, Europe
\endgroup
\pagebreak

@@

The interactions between different plant and pollinator species within a community form complex networks. Macro-ecological analyses of the topology of these networks have revealed common properties across them, such as truncated power-law degree distributions [@jordano1987] or nestedness [@bascompte2003], and has added relevant information of their interaction ecology of plant-pollinator networks across geographic regions [@olesen2002; @traveset2016] and environmental gradients [@ramos2010; @rech2016; @saunders2023]. Although macro-ecological approaches make significant contributions to knowledge, they tend to be rare, and are strongly influenced by the spatio-temporal availability and nature of the data [@burkle2011; @trojelsgaard2013]. For instance, plant-pollinator studies tend to differ in sampling effort and methodology which can impact the structure of the resulting plant-pollinator networks [@gibson2011; @jordano2016]. Further, despite little sampling effort could capture most of the relevant functional species [@hegland2010], most plant-pollinator networks have unobserved interactions because undersampling [@olesen2011; @chacoff2012]. Thus, the lack of strong spatio-temporal coverage along the current intrinsic limitations of sampling plant-pollinator networks, highlight the need of keep gathering and integrating informative species interaction data to properly unravel the different ecological processes at large scales.

Here, we present the EuPPollNet database, an acronym derived from European plant-pollinator networks, which contains harmonized informative interaction data of plants and pollinators at European level. The primary focus of the pollinator taxonomic groups is on the main orders of entomofauna that visit and pollinate flowering plants in Europe. These include insect species from the orders Hymenoptera, Diptera, Lepidoptera, and Coleoptera [@potts2015], which comprise almost the totality of recorded interactions in the EuPPollNet database (`r format(round(main_orders_percent, 2), nsmall = 2)`%). EuPPollNet aims to cover a wide range of taxonomic groups and habitats, while also providing other variables of interest that allow to better control for the ecological context and sampling methods. In addition, EuPPollNet offers a transparent and accessible workflow of its data management and species harmonization that allow to being reused and keep building on it over time. We expect that this database will help to evaluate macro-ecological processes and current gaps of plant-pollinator interactions at European level.

@@

# Data structure

@@

The majority of plant and pollinator species tend to be regionally specific, with only a minor portion of them being shared across a broad range of studies (**Figure 1b**).

## Taxonomy and taxonomic coverage

```{r echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE, out.width="95%", fig.cap= "\\textbf{Figure 2}. Graphs (a-b-c) indicate the accumulation curves for pollinators, plants, and their interactions across sampling sites. Grey solid lines represent 100 randomized accumulation curves, the black solid lines represents the interpolated curve (i.e., the mean across curves), and the red dashed lines illustrates the extrapolated curve for approximately 3000 sampling sites. The solid black points indicate the number of species and interactions contained in the database. Graphs (d-e) indicate the percentage of occurrence (i.e., incidence) of plant and pollinator species across sampling sites. Species on the left (i.e., common) are found in many sampling sites, while species on the right (i.e., rare) are found in few or only a single sampling site. Graph (f) shows the accumulation curve of pollinator species across increasing number of plant species. This last graph uses the same color and shape structure as the ones in the top panel.", fig.width=10, fig.height=6, out.width="115%"}

@@

phylogenetic and geographical biases

@@

The level of statistical difference is indicated with different colours.

@@

@@

#Species per family within SafeNet
bee_spp_safenet = data %>% 
select(Pollinator_rank, Pollinator_accepted_name, Pollinator_order, Pollinator_family)%>%
filter(Pollinator_rank == "SPECIES") %>% 
filter(Pollinator_order == "Hymenoptera") %>% 
filter(!Pollinator_accepted_name == "Apis mellifera") %>% 
filter(Pollinator_family %in% bee_fam) %>% 
group_by(Pollinator_family) %>% 
summarise(Spp_number = n_distinct(Pollinator_accepted_name)) %>% 
mutate(Group = "SafeNet")
#Species per family within Europe
bee_spp_europe =  master_list %>% 
select(Family, Genus, Species, GenusAndSpecies) %>% 
filter(Family %in% bee_fam) %>% 
group_by(Family) %>% 
summarise(Spp_number = n_distinct(GenusAndSpecies)) %>% 
rename(Pollinator_family = Family) %>% 
mutate(Group = "Europe")

@@

p1 = ggplot(bee_fam_coverage, aes(reorder(Pollinator_family,-Spp_number), Spp_number, fill = Group)) +
geom_bar(stat="identity", width=.5, position = "dodge", color = "black") +
coord_cartesian(expand = FALSE) +
xlab(NULL) +
theme_bw() +
theme(axis.text.x = element_text(angle = 45, hjust =1),
      panel.border = element_rect(size=1)) +
ylab("Number of species") +
scale_fill_manual(values = c(viridis(2)))+
ggtitle("Bee family coverage")
#--------------------------------------------------------------------------------------------#
#Continue with syrphids----
#As we don't have subfamilies, let's obtained them from the master list
syrphids = master_list %>% 
filter(Family == "Syrphidae") %>% 
select(Family,Subfamily,  Genus, Species, GenusAndSpecies) %>% 
rename(Pollinator_accepted_name = GenusAndSpecies)
#Syrphid species safenet
syrphid_spp_safenet = data %>% 
select(Pollinator_rank, Pollinator_accepted_name, Pollinator_order, Pollinator_family)%>%
filter(Pollinator_rank == "SPECIES") %>% 
filter(Pollinator_family== "Syrphidae") %>% 
distinct()
#Add subfamily info
syrphid_spp_safenet = left_join(syrphid_spp_safenet, syrphids)
#Summarise spp per subfamily
syrphid_spp_safenet = syrphid_spp_safenet %>% 
group_by(Subfamily) %>% 
summarise(Spp_number = n_distinct(Pollinator_accepted_name)) %>% 
mutate(Group = "SafeNet")
#Get spp per subfamily at european level
syrphid_spp_europe = master_list %>% 
filter(Family == "Syrphidae") %>% 
select(Family, Subfamily, Genus, Species, GenusAndSpecies) %>% 
group_by(Subfamily) %>% 
summarise(Spp_number = n_distinct(GenusAndSpecies)) %>% 
mutate(Group = "Europe")
#Bind both datasets for plotting
syrphid_fam_coverage = bind_rows(syrphid_spp_safenet, syrphid_spp_europe)
#Plot
p2 = ggplot(syrphid_fam_coverage, aes(reorder(Subfamily,-Spp_number), Spp_number, fill = Group)) +
geom_bar(stat="identity", width=.5, position = "dodge", color = "black") +
coord_cartesian(expand = FALSE) +
xlab(NULL) +
ylab(NULL) +
theme_bw() +
theme(axis.text.x = element_text(angle = 45, hjust =1),
      panel.border = element_rect(size=1)) +
scale_fill_manual(values = c(viridis(2)))+
ggtitle("Syrphid subfamily coverage")
#--------------------------------------------------------------------------------------------#
#Continue with butterflies----
#Do it at family level
#Create vector with butterfly families
butt_fam = c(unique(master_list1$family))
#Species per family within SafeNet
butterfly_spp_safenet = data %>% 
select(Pollinator_rank, Pollinator_accepted_name, Pollinator_order, Pollinator_family)%>%
filter(Pollinator_rank == "SPECIES") %>% 
filter(Pollinator_order == "Lepidoptera") %>% 
filter(Pollinator_family %in% butt_fam) %>% 
group_by(Pollinator_family) %>% 
summarise(Spp_number = n_distinct(Pollinator_accepted_name)) %>% 
mutate(Group = "SafeNet")
#Species per family within Europe
butterfly_spp_europe =  master_list1 %>% 
select(family, genus, scientificName) %>% 
group_by(family) %>% 
summarise(Spp_number = n_distinct(scientificName)) %>% 
rename(Pollinator_family = family) %>% 
mutate(Group = "Europe")
#Bind both datasets for plotting
butterfly_fam_coverage = bind_rows(butterfly_spp_safenet, butterfly_spp_europe)
#Plot
p3 = ggplot(butterfly_fam_coverage, aes(reorder(Pollinator_family,-Spp_number), Spp_number, fill = Group)) +
geom_bar(stat="identity", width=.5, position = "dodge", color = "black") +
coord_cartesian(expand = FALSE) +
xlab(NULL) +
ylab(NULL) +
theme_bw() +
theme(axis.text.x = element_text(angle = 45, hjust =1),
      panel.border = element_rect(size=1)) +
scale_fill_manual(values = c(viridis(2)))+
ggtitle("Butterfly family coverage")

#--------------------------------------------------------------------------------------------#
p1+p2+p3 + plot_layout(guides = "collect") & 
theme(plot.title = element_text(size = 14, face = "bold"),
      axis.title.y = element_text(face = "bold"))

>>>>

**Motivation:** Pollinators play a crucial role in maintaining Earth's terrestrial biodiversity and human food production by mediating sexual reproduction for most flowering plants. However, their diversity and role as pollinators are increasingly compromised by rapid human-induced environmental changes. One of the major challenges for pollinator conservation, is the lack of robust generalisable data across space and time to comprehend the conservation status and population trends among different pollinator species. Here, we present the EuPPollNet database, a fully reproducible European-level database containing harmonized taxonomic data on plant-pollinator interactions referenced in both space and time. This database offers an open workflow that allows researchers to track decisions and edit them according to their preferences, while also providing other ecological variables of interest. Furthermore, this work provides an in-depth assessment of the taxonomic and sampling coverage of the database at the European level, complemented by analyses of key structural properties in plant-pollinator networks. We hope this database can help researchers to: 1) identify taxonomic, ecological and geographical gaps of knowledge on plant-pollinator interactions; and 2) explore the impacts of global change to guide future conservation planning for both plant and pollinator species.

@@

**KEYWORDS**\
\begingroup
\small
plant-pollinator networks, plant-pollinator interactions, flowering plants, Angiosperms, pollinators, nestedness, connectance
\endgroup
\pagebreak

@@

The interactions between different plant and pollinator species within a community form complex networks. Macro-ecological analyses of the topology of these networks have revealed common properties across them, such as truncated power-law degree distributions [@jordano1987] or nestedness [@bascompte2003], and has added relevant information of their interaction ecology of plant-pollinator networks across geographic regions [@olesen2002; @traveset2016] and environmental gradients [@ramos2010; @rech2016; @saunders2023]. Although macro-ecological approaches make significant contributions to knowledge, they tend to be rare, and are strongly influenced by the spatio-temporal availability and nature of the data [@burkle2011; @trojelsgaard2013]. For instance, plant-pollinator studies tend to differ in sampling effort and methodology which can impact the structure of the resulting plant-pollinator networks [@gibson2011; @jordano2016]. Further, despite little sampling effort could capture most of the relevant functional species [@hegland2010], most plant-pollinator networks have unobserved interactions because undersampling [@olesen2011; @chacoff2012]. Thus, the lack of strong spatio-temporal coverage along the current intrinsic limitations of sampling plant-pollinator networks, highlight the need of keep gathering and integrating informative species interaction data to properly unravel the different ecological processes that shape plant-pollinator interactions at large scales.

Here, we present the EuPPollNet database, an acronym derived from European plant-pollinator networks, which contains harmonized informative interaction data of plants and pollinators at European level. The primary focus of the pollinator taxonomic groups is on the main orders of entomofauna that visit and pollinate flowering plants in Europe. These include insect species from the orders Hymenoptera, Diptera, Lepidoptera, and Coleoptera [@potts2015], which comprise almost the totality of recorded interactions in the EuPPollNet database (`r format(round(main_orders_percent, 2), nsmall = 2)`%). To understand the scope of the database, we examine the taxonomic and sampling coverage of the different plant and pollinator species at European level with the help of the most up-to-date species checklists and rarefaction analyses. In addition, we investigate how two key structural metrics of plant-pollinator networks, such as connectance and nestedness, change across the latitudinal range of studies and bioclimatic regions. Finally, we also examine whether plant-pollinator networks fulfill the expectation of being more nested than null expectations. EuPPollNet aims to cover a wide range of taxonomic groups and habitats, while also providing other variables of interest that allow to better control for the ecological context and sampling methods. In addition, EuPPollNet offers a transparent and accessible workflow of its data management and species harmonization that allow to being reused and keep building on it over time. We expect that this database will help to evaluate macro-ecological processes and current gaps of plant-pollinator interactions at European level.

@@

## Data structure

@@

```{r echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE}
poll_output = readRDS("../Data/Working_files/pollinator_sampling_coverage.rds")
plant_output = readRDS("../Data/Working_files/plant_sampling_coverage.rds")
int_output = readRDS("../Data/Working_files/interactions_sampling_coverage.rds")
plant_poll_output = readRDS("../Data/Working_files/plant_poll_spp_sampling_coverage.rds")

#Coverage
poll_coverage = poll_output$DataInfo$SC *100
plant_coverage = plant_output$DataInfo$SC *100
int_coverage = int_output$DataInfo$SC *100
plant_poll_coverage = plant_poll_output$DataInfo$SC *100

#Increase doubling effort
##Pollinators
poll_increase = poll_output$iNextEst$Network_id
poll_increase1 = poll_increase %>% filter(method == "observed") %>% pull(qD)
poll_increase2 = poll_increase %>% filter(t == max(t)) %>% pull(qD)
poll_increase3 = (poll_increase2-poll_increase1)/poll_increase1 * 100
##Plants
plant_increase = plant_output$iNextEst$Network_id
plant_increase1 = plant_increase %>% filter(method == "observed") %>% pull(qD)
plant_increase2 = plant_increase %>% filter(t == max(t)) %>% pull(qD)
plant_increase3 = (plant_increase2-plant_increase1)/plant_increase1 * 100
##Interactions
int_increase = int_output$iNextEst$Network_id
int_increase1 = int_increase %>% filter(method == "observed") %>% pull(qD)
int_increase2 = int_increase %>% filter(t == max(t)) %>% pull(qD)
int_increase3 = (int_increase2-int_increase1)/int_increase1 * 100
##Pollinator by plant
plant_poll_increase = plant_poll_output$iNextEst$Plant_id
plant_poll_increase1 = plant_poll_increase %>% filter(method == "observed") %>% pull(qD)
plant_poll_increase2 = plant_poll_increase %>% filter(t == max(t)) %>% pull(qD)
plant_poll_increase3 = (plant_poll_increase2-plant_poll_increase1)/plant_poll_increase1 * 100
```

The estimated sampling coverage of plant and pollinator species within the database across the different sampling sites is approximately `r round(poll_coverage)`% for both taxonomic groups. This suggest that the rarefied accumulation curves of both plant and pollinator species exhibit already a "quasi-asymptotic" growth of species richness by considering the current number of sampling sites or networks (**Figure 2a-2b**). The predicted observed species richness by doubling the sampling effort on the already sampled habitat types within the database will only increase pollinator richness by `r format(round(poll_increase3, 2), nsmall=2)`% and plant richness by `r format(round(plant_increase3, 2), nsmall=2)`%. However, the sampling coverage of interactions is `r format(round(int_coverage, 2), nsmall=2)`%, and by doubling the sampling effort the predicted number of unique interactions recorded will have approximately a twofold increase (`r format(round(int_increase3, 2), nsmall=2)`%; **Figure 2c**). Similarly, to the plant and pollinator species accumulation curves across sampling sites, when we consider the accumulated pollinator richness across sampled plant species, this curve also shows a "quasi-asymptotic" growth with a sampling coverage value of `r format(round(plant_poll_coverage, 2), nsmall=2)`%. The predicted increase of recorded pollinator species by sampling the double of plant species would be `r format(round(plant_poll_increase3, 2), nsmall=2)`%.

Maybe fix last graph? And add it 1st of the last panel (**Figure 2d**)

The majority of plant and pollinator species tend to be regionally specific, with only a minor portion of them being shared across a broad range of studies (**Figures 2e-2f**). For instance, xxxxxxxxx.

The estimated sampling coverage when we consider plants,

```{r echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE, out.width="95%", fig.cap= "\\textbf{Figure 2}. Graphs (a-b-c) indicate the accumulation curves for pollinators, plants, and their interactions across sampling sites. Grey solid lines represent 100 randomized accumulation curves, the black solid lines represents the interpolated curve (i.e., the mean across curves), and the red dashed lines illustrates the extrapolated curve for approximately 3000 sampling sites. The solid black points indicate the number of species and interactions contained in the database. Graph (d) shows the accumulation curve of pollinator species across increasing number of plant species. This last graph uses the same color and shape structure as the ones in the top panel. Graphs (e-f) indicate the percentage of occurrence (i.e., incidence) of plant and pollinator species across sampling sites. Species on the left (i.e., common) are found in many sampling sites, while species on the right (i.e., rare) are found in few or only a single sampling site.", fig.width=10, fig.height=6, out.width="115%"}

@@

The database contains xxxx species of pollinators and xxxx species of plants. Most species of pollinators are Hymenoptera species (%), and most of these species are bees (%).

## Network properties

@@

@@

Although Europe...

@@

#Species per family within SafeNet
bee_spp_safenet = data %>% 
select(Pollinator_rank, Pollinator_accepted_name, Pollinator_order, Pollinator_family)%>%
filter(Pollinator_rank == "SPECIES") %>% 
filter(Pollinator_order == "Hymenoptera") %>% 
filter(!Pollinator_accepted_name == "Apis mellifera") %>% 
filter(Pollinator_family %in% bee_fam) %>% 
group_by(Pollinator_family) %>% 
summarise(Spp_number = n_distinct(Pollinator_accepted_name)) %>% 
mutate(Group = "EuPPollNet")
#Species per family within Europe
bee_spp_europe =  master_list %>% 
select(Family, Genus, Species, GenusAndSpecies) %>% 
filter(Family %in% bee_fam) %>% 
group_by(Family) %>% 
summarise(Spp_number = n_distinct(GenusAndSpecies)) %>% 
rename(Pollinator_family = Family) %>% 
mutate(Group = "Europe")

@@

p1 = ggplot(bee_fam_coverage, aes(reorder(Pollinator_family,-Spp_number), Spp_number, fill = Group)) +
geom_bar(stat="identity", width=.5, position = "dodge", color = "black") +
coord_cartesian(expand = FALSE) +
xlab(NULL) +
theme_bw() +
theme(axis.text.x = element_text(angle = 45, hjust =1),
      panel.border = element_rect(size=1)) +
ylab("Number of species") +
scale_fill_manual(values = c("tan2", "lightgrey"))+
ggtitle("Bee family coverage")
#--------------------------------------------------------------------------------------------#
#Continue with syrphids----
#As we don't have subfamilies, let's obtained them from the master list
syrphids = master_list %>% 
filter(Family == "Syrphidae") %>% 
select(Family,Subfamily,  Genus, Species, GenusAndSpecies) %>% 
rename(Pollinator_accepted_name = GenusAndSpecies)
#Syrphid species safenet
syrphid_spp_safenet = data %>% 
select(Pollinator_rank, Pollinator_accepted_name, Pollinator_order, Pollinator_family)%>%
filter(Pollinator_rank == "SPECIES") %>% 
filter(Pollinator_family== "Syrphidae") %>% 
distinct()
#Add subfamily info
syrphid_spp_safenet = left_join(syrphid_spp_safenet, syrphids)
#Summarise spp per subfamily
syrphid_spp_safenet = syrphid_spp_safenet %>% 
group_by(Subfamily) %>% 
summarise(Spp_number = n_distinct(Pollinator_accepted_name)) %>% 
mutate(Group = "EuPPollNet")
#Get spp per subfamily at european level
syrphid_spp_europe = master_list %>% 
filter(Family == "Syrphidae") %>% 
select(Family, Subfamily, Genus, Species, GenusAndSpecies) %>% 
group_by(Subfamily) %>% 
summarise(Spp_number = n_distinct(GenusAndSpecies)) %>% 
mutate(Group = "Europe")
#Bind both datasets for plotting
syrphid_fam_coverage = bind_rows(syrphid_spp_safenet, syrphid_spp_europe)
#Plot
p2 = ggplot(syrphid_fam_coverage, aes(reorder(Subfamily,-Spp_number), Spp_number, fill = Group)) +
geom_bar(stat="identity", width=.5, position = "dodge", color = "black") +
coord_cartesian(expand = FALSE) +
xlab(NULL) +
ylab(NULL) +
theme_bw() +
theme(axis.text.x = element_text(angle = 45, hjust =1),
      panel.border = element_rect(size=1)) +
scale_fill_manual(values = c("tan2", "lightgrey"))+
ggtitle("Syrphid subfamily coverage")
#--------------------------------------------------------------------------------------------#
#Continue with butterflies----
#Do it at family level
#Create vector with butterfly families
butt_fam = c(unique(master_list1$family))
#Species per family within SafeNet
butterfly_spp_safenet = data %>% 
select(Pollinator_rank, Pollinator_accepted_name, Pollinator_order, Pollinator_family)%>%
filter(Pollinator_rank == "SPECIES") %>% 
filter(Pollinator_order == "Lepidoptera") %>% 
filter(Pollinator_family %in% butt_fam) %>% 
group_by(Pollinator_family) %>% 
summarise(Spp_number = n_distinct(Pollinator_accepted_name)) %>% 
mutate(Group = "EuPPollNet")
#Species per family within Europe
butterfly_spp_europe =  master_list1 %>% 
select(family, genus, scientificName) %>% 
group_by(family) %>% 
summarise(Spp_number = n_distinct(scientificName)) %>% 
rename(Pollinator_family = family) %>% 
mutate(Group = "Europe")
#Bind both datasets for plotting
butterfly_fam_coverage = bind_rows(butterfly_spp_safenet, butterfly_spp_europe)
#Plot
p3 = ggplot(butterfly_fam_coverage, aes(reorder(Pollinator_family,-Spp_number), Spp_number, fill = Group)) +
geom_bar(stat="identity", width=.5, position = "dodge", color = "black") +
coord_cartesian(expand = FALSE) +
xlab(NULL) +
ylab(NULL) +
theme_bw() +
theme(axis.text.x = element_text(angle = 45, hjust =1),
      panel.border = element_rect(size=1)) +
scale_fill_manual(values = c("tan2", "lightgrey"))+
ggtitle("Butterfly family coverage")

#--------------------------------------------------------------------------------------------#
p1+p2+p3 + plot_layout(guides = "collect") & 
theme(plot.title = element_text(size = 13, face = "bold"),
      axis.title.y = element_text(face = "bold"))
